Single Bivalent Level Analysis

### 2020-01-23 doc/Bivalent_Analysis.Rmd
updates can knit!

```{r long.chrm.dataset, echo=FALSE, eval=FALSE}

#how many biv's are from cells with >15 bivalents

al.15_bivs <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]#150 cells have al 15

long_BivData_set <- Curated_BivData[ Curated_BivData$fileName %in% al.15_bivs$fileName, ] #across those 150 cells are 2654 biv measures

#table(potential.data.set$category)
long_BivData_set$fileName <- as.character(long_BivData_set$fileName)
Biv_Data_mean_table$fileName <- as.character(Biv_Data_mean_table$fileName)

#pull out the biv's which exceed the cell's Q4
#for each line of long_BivData_set
# check if it's SC length is greater than the Q4 for the given cell

long_BivData_set99 <- long_BivData_set

for(  i in 1:length( long_BivData_set99$fileName)  ){
  long_BivData_set99$cell.Q4[i] <- Biv_Data_mean_table[ which(Biv_Data_mean_table$fileName == long_BivData_set99$fileName[i]), ]
}

for(  k in 1:length( long_BivData_set99$fileName)  ){
long_BivData_set99$above.Q4[k] <- ifelse( (long_BivData_set99$cell.Q4 < long_BivData_set99$chromosomeLength), "yes", "no")
}
#This isn't working either

#for(i in 1:length( long_BivData_set$fileName)  ){ 
  # print(i)
#  current_obj_fileName <- long_BivData_set$fileName[i]
#  current_obj_SC_length <- long_BivData_set$chromosomeLength[i]  
#  current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == current_obj_fileName) ]
# current.Q4 <- Biv_Data_mean_table$Qar.4[which(Biv_Data_mean_table$fileName == long_BivData_set$fileName[i]) ]
#  long_BivData_set$above.Q4 <-ifelse( (current.Q4 < current_obj_SC_length), 1, 0)
#  print( paste( c(current.Q4, current_obj_fileName, current_obj_SC_length), sep = "_" ) )
#}
#this loop isn't working

@@

The dataset is unbalanced in some ways -- integrating manual measures might fix. MSM under rep, since most cell had high overlap.

>>>>

Biv_Data_mean_table has the Q4 values

```{r long.chrm.dataset, echo=FALSE}

#1. how many biv's are from cells with >15 bivalents

over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

#check for blank obj.id
BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")

#check for dup
length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]

#remove duplicate obj.ids
BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]

#make sure obj.id and fileName are charcter for str stuff
BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)

#make sure 
missing_from_cell_table <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% Biv_Data_mean_table$fileName, ]#there are 75, objs 3 cells
#these are bivalents which don't appear the in cell level table, which can break the loop
BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]

#final cleaned BivData table, 2428

for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){

  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_")[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

  the.Q4 <- Biv_Data_mean_table$Qar.4[which( grepl(new.egg, Biv_Data_mean_table$fileName ) )]
  #the above can produe mulitple Q4

  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
  #these are just using the last index,

  the.FN <- Biv_Data_mean_table$fileName[which(grepl(new.egg, Biv_Data_mean_table$fileName ) )]

  BivData_from_15biv_cells$Q4.FN[i] <- the.FN
   # print(c(i, the.FN, the.Q4))

BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= BivData_from_15biv_cells$cell.Q4), "yes", "no" )

real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]
table(real.long.bivData$fileName)

#!!! omg it finally worked!~!!#OMFG

@@

Well fixing those bigs took forever

```{r show.long.SC, echo=FALSE}

female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))
female_long_biv_plot

male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))
male_long_biv_plot

These plots show the SC lengths for bivalents I'm calling the 'long SC data set'. They are supposed to be the longest 4-5 SC from cells where I could get good measures. These longer bivalents are useful because their patterns shouldn't be affect by chromsome size effect (which effects, CO position). Hopefully this data set will have less noise from chromosome identity, but there was still data missing (they don't come from whole cell measures).
